---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hierarchical

`hierarchical` provides tools for manipulations, transformations and analysis of hierarchical data


## Example of decomposition of financial ratios

### Simulate random data

```{r}
library(data.table)
library(data.tree)

# load example data
data("Angell", package = "carData")
DT <- as.data.table(Angell, keep.rownames = TRUE)[sample(1:.N, 20)]
setnames(DT, "rn", "name")

# simulate financial data
DT <- rbindlist(list("2016" = DT, "2007" = DT), idcol = "year")
DT[, cost   := 1000*rgamma(.N, 30, 30)]
DT[, income := 1000*rgamma(.N, 31, 31)]

# compute first differences
setkey(DT, region, name, year)
DT[, cost_diff   := c(NA, diff(cost)),   by = .(region, name)]
DT[, income_diff := c(NA, diff(income)), by = .(region, name)]
DT[, cost_to_income := cost / income]
```

### Convertion of data.table into a data.tree
```{r main}
library(hierarchical)

tree <- dt2tree_ratio(DT, 
                      tree_name       = "Total", 
                      numerator_name  = "cost", 
                      denomiator_name = "income", 
                      dim_names       = c("region", "name"))
```


### Decomposition of ratios

```{r compute decomp, fig.height=7, fig.width=7}
tree$Do(decomp_ratios_root_full, 
        root_denominator_curr = tree$denominator_curr,
        root_denominator_lag  = tree$denominator_lag, 
        multipl = 1)
```

### Output

You can print "ratio_effect", "weight_effect" and "residual_effect" using `data.tree` methods:

```{r decomp table}
print(tree, "ratio_curr", "ratio_effect", "weight_effect")
```


...or print it using `ggplot2`

```{r decomp plot, fig.height=6, fig.width=7, dpi=96, out.width='70%'}
library(ggplot2)
plot_tree_ratio(tree)  +
  ggtitle("Decompositon of consolidated cost-to-income ratio") +
  xlab("Entitity") + 
  theme(legend.text = element_text(size = 7))
```


